---
title: "MachineLearning"
author: "Meagan Marziani"
date: "Thursday, May 21, 2015"
output: pdf_document
---

##Practical Machine Learning 

Load the training data and specify the NAs in our data
```{r}
trainset <- read.csv("C:/Users/Meagan/Desktop/Coursera/MachineLearningProject/pml_training.csv", na.strings=c("NA","#DIV/0!", ""))
testset <- read.csv('C:/Users/Meagan/Desktop/Coursera/MachineLearningProject/pml_testing.csv', na.strings=c("NA","#DIV/0!", ""))
```
Load the necessary packages
```{r}
library(caret)
library(randomForest)
```
Basic exploratory data analysis of our train set
```{r}
dim(trainset)
table(trainset$classe)
plot(trainset$classe, col="black", main="Levels of the variable classe", 
     xlab="classe levels", ylab="Frequency")
```
Basic exploratory data analysis of our test set
```{r}
dim(testset)
```
We will predict the classe for this set

Set the seed so our numbers don't change
```{r}
set.seed(1892)
```
Delete unecessary columns 
```{r}
trainset <-trainset[,colSums(is.na(trainset)) == 0]
testset <-testset[,colSums(is.na(testset)) == 0]

trainset   <-trainset[,-c(1:7)]
testset <-testset[,-c(1:7)]
```
See new dimensions
```{r}
dim(trainset)
dim(testset)
```
In order to perform cross-validation, the training data set is partionned 
into 2 sets: training (75%) and testing (25%).
This will be performed using random subsampling without replacement.
```{r}
subsamp <- createDataPartition(trainset$classe, p=0.75, list=FALSE)
subtrain <- trainset[subsamp, ]
subtest <-   trainset[-subsamp, ]

model <- randomForest(classe ~., data=subtrain, method="class")
prediction <- predict(model, subtest, type="class")
confusionMatrix(prediction, subtest$classe)

Accuracy <- 0.9953
SampleError <- 1-Accuracy
SampleError
```

My out of sample error is expected to be 0.0047 or 0.4%.  

Save the test answers into individual text files for submission. 
```{r}
setwd("C:/Users/Meagan/Desktop/Coursera/MachineLearningProject/20_tests")

predictfinal <- predict(model, testset, type="class")
predictfinal

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictfinal)
```








